---
title: "Assignment Week 6"
author: "Leonard C. Steg"
date: '2022-04-01'
output: html_document
---

```{r, eval = F}
BiocManager::install(c("MotifDb", "universalmotif", "TFBSTools", "memes", "PWMEnrich"))
```


```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(epiwraps)
  library(MotifDb)
  library(universalmotif)
  library(TFBSTools)
  library(memes)
  library(PWMEnrich)
})
```
### Download Nanog ChIP peaks

```{r, eval=FALSE}
dir.create("EncodePeaks")
download.file("https://www.encodeproject.org/files/ENCFF097ISB/@@download/ENCFF097ISB.bed.gz", destfile = "/mnt/groupMansuy/leo/epibioinfo/week06/EncodePeaks/nanog.bed.gz")
```

### Load and prepare the peaks for the analysis
```{r}
peaks <- rtracklayer::import("/mnt/groupMansuy/leo/epibioinfo/week06/EncodePeaks/nanog.bed.gz", format = "narrowPeak")
seqlevelsStyle(peaks) <- "Ensembl"
peak_centers <- resize(peaks, fix="center", width=100)
```

### Load genome and identify sequences of peaks
```{r}
ah <- AnnotationHub()
genome <- ah[["AH49775"]]

peak_seqs <- memes::get_sequence(peak_centers, genome)
```

```{r}
motifs <- query(MotifDb, "Nanog")
names(motifs)
motif <- motifs[["Mmusculus-HOCOMOCOv10-NANOG_MOUSE.H10MO.A"]]
view_motifs(motif)
```


## Scanning for motif in peaks and answering question 1
```{r}
moi <- memes::runFimo(peak_seqs,  convert_motifs(motif), meme_path = "/mnt/IM/conda/bin/")

length(peaks)
sum(overlapsAny(peaks, moi))
sum(overlapsAny(peaks, moi))/length(peaks)
```

From all Nanog peaks (n = 16037), 4311 (26.88%) contain the chosen Nanog motif


## Scanning for motif in whole genome and answering question 2
### Loading in genome and running runFimo
```{r}
genome <- import(genome)
moi_genome <- memes::runFimo(genome, convert_motifs(motif), meme_path = "/mnt/IM/conda/bin")

length(moi_genome)
sum(overlapsAny(moi_genome, peaks))
sum(overlapsAny(moi_genome, peaks))/length(moi_genome)
```

From all motifs in the genome (n = 2811800), 10785 (0.38%) are bound by Nanog (have a peak)